Approximate Building Blocks for Image Parsing and Synthesis

### 2014-11-05 Paper/paper.tex
paper update

\textbf{Texture synthesis} has received a lot of attention in computer graphics, see \cite{Wei2009STAR} for a comprehensive survey. In order to overcome the limitations of the Markov Random Field (MRF) model, various forms of guidance have been employed, for example, using an additional image layer with user annotations \cite{Hertzmann2001IA} or on-the-fly user interaction ~\cite{Barnes2009,Hu2013PPI}. Automating the creation of guidance map is still an on-going research topics, with some success in low level feature maps. For example, Lefebvre and Hoppe~\shortcite{Lefebvre2006ATS} use distance transforms on feature maps to guide the synthesis of high-quality textures with pronounced meso-structure. However, the feature transform cannot structure complex images beyond textures. Rosenberger et al.~\shortcite{Rosenberger2009LSS} extend the idea to irregular textures with non-stationary statistics, such as aged or weathered surfaces. Dishler et al.~\shortcite{Dischler2002} use ``texture particles'' as landmarks, similar to our guidance scheme; however, only textures (with some meso-structure) are synthesized due to limitations in the detection stage. Tiling grammars similar to our higher-level structure have been recently employed to guide the placement of discrete elements~\cite{Ma2011DET,Ma2013DET}, however our method handles image data rather than discrete elements and builds the tiling grammar fully automatically from an image. Our application is also related to image retargeting, recomposition and inpainting. Along this direction, various methods have been experimented to avoid local optimal solutions results from the MRFs model. For example, \cite{Simakov2008SV} uses bidirectional matching and a gradual resizing procedure for retain the image structure in retargeting (especially for image shrinking); \cite{Pritch09ICCV,He2012PO} studied pixel offset, and use graph cut to find the optimal pixel assignment for the output image that produces a smooth offset field. Nonetheless, these methods all operate on image pixels, hence can not eliminate structural artefacts. An exception is \cite{Wu2010SS}, which use deformable lattice to retain structure during image resizing. However their method is restricted to images has a global repetition (such as lattice), and is difficult to be used for recomposition. In this paper we will show using additional structure knowledge is able to improve the performance of the conventional pixel based MRF methods.

@@

To overcome this problem, we incorporate building block constraints into the energy function and redefine the smoothness term as the following:

@@

Here $E_{s}^{pixel}$ is the conventional pixel-based smoothness term as used by \cite{Pritch09ICCV} and \cite{He2012PO}. It penalizes two neighboring labels $a, b$ that creates a seam for adjacent pixels $x, x'$:
%
\begin{equation}
E_{s}^{pixel}(x, x', a, b) = |I(x + s_{a}) - I(x + s_{b})| + |I(x' + s_{a}) - I(x' + s_{b})|
\label{eq:ESmooth}
\end{equation}
%
In addition to it, we define two building block costs: the class smoothness cost $E_{s}^{class}$ and the position smoothness cost $E_{s}^{position}$. They are Potts models with different distance metrics: The class smoothness term uses the hamming distance between building block class labels \revised{$E_{s}^{class} = $}, so is able to penalizes falsely stitched building block classes. The position smoothness cost uses Eculidean distance between pixels' positions in their parent building block's local frame \revised{$E_{s}^{position} = $}, hence penalizes internally mis-configured building blocks. Notice this term is only activated for pixels that are assigned with the same class label, and the position is normalized according to the size of the parent building block. In this way the position smoothness term has the class smoothness cost as the upper bound. Only in this way the sub-modular constraint for graph cut optimization \cite{BOYKOVpami2001} can be met.

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.95\linewidth]{Fig/IntermediumSyn/IntermediumSyn}
  \caption{IntermediumSyn of each step.} \label{fig:IntermediumSyn}
\end{figure}

\revised{Notice the theoretical upper bound of these three costs are different: $sup E_{s}^{pixel} = 255 \times 2 = 500$, $sup E_{s}^{class} = 1 \times 2 = 2$, and $sup E_{s}^{position} = 1 \times 4 = 4$. In practice we should weight these costs properly so none of them, especially $E_{s}^{pixel}$, becomes predominant. However, our observation is that it is rare to have large $E_{s}^{pixel}$ surviving in the result -- as they will be naturally removed as clear seams. In practice the ``semantic'' artifacts are caused by pixels of similar values, in which case $E_{s}^{pixel}$ is usually less than 50. Based on this observation, we set $(\omega_{1}, \omega_{2}, \omega_{3})$ to $(1, 10, 2)$ to remove artifacts that couldn't not be handled by $E_{s}^{pixel}$ alone. Figure \ref{fig:IntermediumSyn} b) shows the retargeting result of a) using \cite{Pritch09ICCV}. Although the stitching of the arches is clearly mis-stitched to human, such artifacts can not be avoided if only pixel smoothness is considered. In cooperating the class smoothness cost and the position smoothness cost is able to reduce the artifacts (Figure \ref{fig:IntermediumSyn} c)). The improvement is more obvious in the class label map, shown in the bottom row of the figure. Especially, the noisy yellow labels at the floor of b) is removed, which results in more plausible synthesis in c).}

\subsubsection{Building Blocks Based Offsets}
\revised{
As pointed out in \cite{He2012PO}, the selection of offset labels is crucially important to stitching performance. Regular sampling has a risk of being too dense or too sparse for an optimal solution. For example, the artifacts in the middle part of the Figure \ref{fig:IntermediumSyn} c) (glasses window) is caused by a local optimal solution due to offset samples that do not conform to the regularity in the facade. In practice we found synthesis can be significantly improved by using offset labels generated from building blocks statistics, as shown in Figure \ref{fig:IntermediumSyn} d). Our building blocks based offsets sampling is performed as following: we first compute a probability density map (PDF) of offset statistics by accumulating pairwise building blocks offsets. Only building blocks of the same class can vote for their offset. We then detection local peaks in the PDF as \cite{He2012PO} did. These local peaks can be directly used as candidate offsets for image completion, as described in \cite{He2012PO}. For our task of image retargeting, a larger montage must be created so the synthesis domain can be fully covered. In theory this can be achieved by recursively sampling new peaks from the PDF. However the number of offsets will increase as a power function and the search of offset labeling quickly becomes infeasible in practice. To handle this problem, we find predominant offsets and use them as grid generators to efficiently create the montage: the first generator is the strongest peak in the PDF, and the second generator is the next strongest peak that satisfies a minimum angular distance to the first generator. The minimum angular (set to $\frac{\pi}{3}$) requirement de-correlates the two generators so the montage can be efficiently expanded as a 2-D grid in both horizontal and vertical directions. Additional generators can be combined with these two to create extra grids in the pursuit of higher variability in the result. In practice, we find a 2d grid is a good trade off between stability and variability.}

\revised{Since we use a grid based offset layout, it is important to investigate whether such layout can handle various size of the retargeting image. Conventional pixel based method \cite{Pritch09ICCV}\cite{He2012PO} often generates mis-stitched or mis-aligned objects when the image size that is a fraction of the generator (top row, Figure \ref{fig:incompatible}). Figure \ref{fig:incompatible} a) and b) (top row) are created using \cite{He2012PO}, which mis-aligned objects to accommodate them in the retargeted image. Figure \ref{fig:incompatible} c) (top row) is created using \cite{Pritch09ICCV}, which created artificial stitching that distorts the shape of the windows. Such artifacts can be better revealed in the building block label layer, as shown next to the retargeted images. In contrast, building blocks give additional constraints that is able to correct some of these artifacts (bottom row, Figure \ref{fig:incompatible}). Specifically, in Figure \ref{fig:incompatible} a) (bottom), our method consistently uses the ornament (green building block) to fill the middle of the facade and in Figure \ref{fig:incompatible} b) (bottom), it snaps balcony and the lower floor window (yellow and blue building blocks). Notice for these cases, we use the same generators of the top row to create the results in the bottom row, so the improvement purely comes from the building blocks constraints. However, we can also create a montage that is larger than the retarget image and allow regular structure to propagate beyond the image boundary. This is used to create Figure \ref{fig:incompatible} c) (bottom). Despite these successful examples, it is worth mentioning failure cases can also be created for strongly incompatible image sizes. This will be discussed together with other failure cases later in the paper.}

\begin{figure*}[t!]
  \centering
  \includegraphics[width=0.95\linewidth]{Fig/incompatible/incompatible}
  \caption{incompatible.} \label{fig:incompatible}
\end{figure*}

@@

Similar to \cite{He2012PO}, we use co-occurrence statistics to generate candidate labels. This improves the quality and efficiency of the synthesis. Different to \cite{He2012PO}, we do not need to compute pairwise pixel distance to get the offset image statistics. Instead, can we directly use the co-occurrence statistics between building blocks. In practice we can sample candidate labels on the fly to generate randomness in the synthesis results (see our supplementary video for an example).

@@

\textbf{Limitations and Future Work:} Our method has a number of limitations, some of which could be addressed in future work. First of all, it is limited to translational building blocks. While the feature and learning pipeline is able to compensate some distortions, strong perspective, scaling, or rotation cannot be handled. We believe that this can be extended quite easily; the main challenge is that we need to replace the global alignment of co-occurrence maps with local alignments that take the coordinate frame of pairwise matches into account. Second, as shown in the pressure test, statistics favors the grid-based detection when the demand of shape accuracy is high. This implies higher level regularity, when observed, should be incorporated to improve our detection. This is important because only high quality detection is able to help later applications. For example, inaccurate detection can only downgrades image synthesis by imposing useless constraints. Further, although our model uses tiling grammar to extend the local MRF model, the inference of such grammar is combinatorially hard. In many cases our model still fail to find a global optimal solution and consequently, failed to correct structural artefact in the synthesized image. Here, we could think of including more comprehensive, global relations such as symmetry or hierarchy~\cite{Hu2013PPI}, or we could use the inverse procedural modeling approach in the spirit of \cite{BOKELOHsig2010,Sylvain2010AT} to obtain stronger guarantees for a subset of structures. In summary, conventional pixel-level synthesis has to rely on MRF-based texture generation; while guidance helps, we still observe artifacts and failure cases. A more fundamental understanding of image structure would be necessary to obtain more powerful models.

>>>>

\textbf{Texture synthesis} has received a lot of attention in computer graphics, see \cite{Wei2009STAR} for a comprehensive survey. In order to overcome the limitations of the Markov Random Field (MRF) model, various forms of guidance have been employed, for example, using an additional image layer with user annotations \cite{Hertzmann2001IA} or on-the-fly user interaction ~\cite{Barnes2009,Hu2013PPI}. Automating the creation of guidance map is still an on-going research topics, with some success in low level feature maps. For example, Lefebvre and Hoppe~\shortcite{Lefebvre2006ATS} use distance transforms on feature maps to guide the synthesis of high-quality textures with pronounced meso-structure. However, the feature transform cannot structure complex images beyond textures. Rosenberger et al.~\shortcite{Rosenberger2009LSS} extend the idea to irregular textures with non-stationary statistics, such as aged or weathered surfaces. Dishler et al.~\shortcite{Dischler2002} use ``texture particles'' as landmarks, similar to our guidance scheme; however, only textures (with some meso-structure) are synthesized due to limitations in the detection stage. Tiling grammars have been recently employed to guide the placement of discrete elements~\cite{Ma2011DET,Ma2013DET}, however our method handles image data, such as pixels color and class labels, rather than discrete elements. Our application is also related to image retargeting, recomposition and inpainting. Along this direction, various methods have been experimented to avoid local optimal solutions results from the MRFs model. For example, \cite{Simakov2008SV} uses bidirectional matching and a gradual resizing procedure for retain the image structure in retargeting (especially for image shrinking); \cite{Pritch09ICCV,He2012PO} studied pixel offset, and use graph cut to find the optimal pixel assignment for the output image that produces a smooth offset field. Nonetheless, these methods all operate on image pixels, hence can not eliminate structural artefacts. An exception is \cite{Wu2010SS}, which use deformable lattice to retain structure during image resizing. However their method is restricted to images has a global repetition (such as lattice), and is difficult to be used for recomposition. In this paper we will show using additional structure knowledge is able to improve the performance of the conventional pixel based MRF methods.

@@

To overcome this problem, we incorporate building block constraints into the energy function. \revised{The idea is straight forward: as frequently appearing objects, building blocks should be prevented from mis-configuration and distortion.} Based on this idea, we redefine the smoothness term as the following:

@@

Here $E_{s}^{pixel}$ is the conventional pixel-based smoothness term as used by \cite{Pritch09ICCV} and \cite{He2012PO}. It penalizes two neighboring labels $a, b$ that creates a seam for adjacent pixels $x, x'$:
%
\begin{equation}
E_{s}^{pixel}(x, x', a, b) = |I(x + s_{a}) - I(x + s_{b})| + |I(x' + s_{a}) - I(x' + s_{b})|
\label{eq:ESmooth}
\end{equation}
%
In addition to it, we define two building block costs: the class smoothness cost $E_{s}^{class}$ and the position smoothness cost $E_{s}^{position}$. They are Potts models with different properties.

\revised{\textbf{Class smoothness term:} Let $l(x)$ be the building blocks class label for pixel $x$, the class smoothness term computes the hamming distance between class labels: $E_{s}^{class} = l(x + s_{a}) \neq l(x + s_{b}) + l(x' + s_{a}) \neq l(x' + s_{b})$.} It penalizes falsely stitching together pixels from different building block classes \revised{so is able to reduce mis-configurations}.

\revised{\textbf{Position smoothness term:} Let $(r(x), c(x))$ be the relative row and column position for pixel $x$ in its parent building blocks's local frame. The position smoothness cost computes the Manhattan distance between the relative positions: $E_{s}^{position} = w(x)(|r(x + s_{a}) - r(x + s_{b})| + |c(x + s_{a}) - c(x + s_{b})|) + w(x')(|r(x' + s_{a}) - r(x' + s_{b})| + |c(x' + s_{a}) - c(x' + s_{b})|)$}. It penalizes internally mis-configured building blocks \revised{so is able to prevent distortions}. Notice this term is only activated for pixels that are assigned with the same class label, \revised{that is $w(x) = l(x + s_{a}) == l(x + s_{b})$. In the meantime $r(x)$ and $c(x)$ are normalized according to the size of the parent building block so $r(x) \in [0, 1]$ and $c(x) \in [0, 1]$}. In this way the position smoothness term has the class smoothness cost as the upper bound. Only in this way the sub-modular constraint for graph cut optimization \cite{BOYKOVpami2001} can be met.

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.95\linewidth]{Fig/IntermediumSyn/IntermediumSyn}
  \caption{\revised{Compared to \protect\cite{Pritch09ICCV}, using building blocks is able to reduce the artifacts in the retargeting c). The result is further improved using building blocks offset statistics d).}} \label{fig:IntermediumSyn}
\end{figure}

 \revised{\textbf{Weights:} Notice the three costs in Equation \ref{eq:ESmooth} share the same lower bound (zero) but different upper bounds: $sup E_{s}^{pixel} = 255 \times 2 = 500$, $sup E_{s}^{class} = 1 \times 2 = 2$, and $sup E_{s}^{position} = 1 \times 4 = 4$. This implies we should weight them carefully so none of them, especially $E_{s}^{pixel}$, becomes predominant. In practice we also observed that artifacts in the conventional MRF models are rarely created by large $E_{s}^{pixel}$, as large pixel costs are naturally removed by graph cut as obvious seams. In fact, these artifacts are often created by stitching together pixels of similar values, in which case $E_{s}^{pixel}$ is small. This leads to the conclusion that building blocks costs should give enough influence, but not overkill, the low pixel costs. In practice we find $\omega_{1} = 1, \omega_{2} = 10, \omega_{3} = 2$ works robustly across a wide range of images. Although they can be further tuned on a per image basis for better results, we fix them to these default values in the later user study for fair comparisons.}

\revised{Figure \ref{fig:IntermediumSyn} shows an example of how the building blocks costs improve the result. The conventional MRF based method \cite{Pritch09ICCV} creates the result in Figure \ref{fig:IntermediumSyn} b). While the stitching of the arches is clearly artificial to human, it is hard to be avoided if only pixel cost is considered. With the help of the building blocks costs we are able to reduce the artifacts to Figure \ref{fig:IntermediumSyn} c). The improvement is more obvious in the class label map, shown in the bottom row of the figure. Especially, the noisy yellow labels at the floor of b) is removed, which results in more plausible synthesis in c).}

\subsubsection{Building Blocks Based Offsets}
\revised{Another important factor for achieving high quality synthesis results is the input offset labels for graph cut optimization. These labels directly decide the complexity of the search space and whether the minimized energy leads to a visually plausible image. For image completion, regular sampled offsets can be too dense or too sparse for achieving high quality results, as observed by \cite{He2012PO}. This is also true for image regarting. For example, in Figure \ref{fig:IntermediumSyn} c) a window is squeezed at the center of the image, despite the use of building blocks constraints. This is due to the fact that offset labels are sampled using a grid generator that deviates from regularities in the facade. In summary, building blocks constraints are less useful if the search space does not accommodate a useful solution in the first place.}

\revised{To address this problem, we sample offset labels from building blocks offset statistics. In this way both the search space and the constraints are based on the knowledge of building blocks. This often leads to further improvement of the synthesis results, as shown in Figure \ref{fig:IntermediumSyn} d). Our sampling contains two main steps: It first computes predominant offsets between the same class of building blocks. These offsets are then used as generators to regularly generate input labels for graph cut. Specifically, it works as follows: A probability density map (PDF) is computed by accumulating votes of pairwise building blocks offsets. Only building blocks of the same class are considered in the voting. We then detect local peaks in the PDF using non-maximum suppression. These local peaks were used in previous works as offset labels for image completion, as described in \cite{He2012PO}. Differently, for our task of image retargeting a larger montage must be created to fully cover the synthesis domain. In theory this can be achieved by recursively sampling new peaks from the PDF. However the number of offsets will be increased by a power function and the search becomes infeasible in practice. To handle this problem, we find predominant offsets and use them as grid generators to efficiently create the montage: the first generator is the strongest peak in the PDF, and the second generator is the next strongest peak that satisfies a minimum angular distance to the first generator. The minimum angular (set to $\frac{\pi}{3}$) requirement de-correlates the two generators so the montage can be efficiently expanded as a 2-D grid in both horizontal and vertical directions. In the pursuit of higher variability in the result, additional generators can be combined with these two to create extra grids. For example, randomness in the synthesis results can be created using a near-regular grid. However, in practice we found a regular grid spanned by two predominate generators trades off well between stability and variability.}

\revised{Since we use grid based offsets, it is important to investigate whether our method can handle various size of the retargeting image, especially when the image size is a fraction of the generators. In this case, conventional pixel based method \cite{Pritch09ICCV}\cite{He2012PO} often generates mis-stitched or mis-aligned objects: Figure \ref{fig:incompatible} a) and b) (top) are created using \cite{He2012PO}. In these examples objects are mis-aligned so they can be fit to the retargeted image. These artifacts can be better revealed in the building block label maps, as shown next to the retargeted images. In contrast, our method (bottom) is able to correct these artifacts using the same input offset labels. Specifically, in Figure \ref{fig:incompatible} a), our method consistently uses the ornament (green building block) to fill the middle of the facade; in Figure \ref{fig:incompatible} b), it snaps balcony and the lower floor window (yellow and blue building blocks). Since the same offset labels are used to create the results, these improvement purely comes from the building blocks constraints (Equation \ref{eq:EPixelSmooth}). Figure \ref{fig:incompatible} c) (top) is created using \cite{Pritch09ICCV} with a montange that is of the same size as the retargeted image. In this case the windows are distorted. Since we use a montage that is larger than the retarget image, regular structures are allowed to propagate beyond the image boundary. This is used to create Figure \ref{fig:incompatible} c) (bottom). Despite these successful examples, it is worth mentioning failure cases can occur for strongly incompatible image sizes. This will be discussed later in the paper with other failure cases.}

\begin{figure*}[t!]
  \centering
  \includegraphics[width=0.95\linewidth]{Fig/incompatible/incompatible}
  \caption{\revised{Our method is able to reduce artifacts when the synthesis image is of sizes that are incompatible to the offset labels. Top: result from \protect\cite{He2012PO} and \protect\cite{Pritch09ICCV}. Bottom row: our results.}} \label{fig:incompatible}
\end{figure*}

@@

@@

\textbf{Limitations and Future Work:} Our method has a number of limitations, some of which could be addressed in future work. First of all, it is limited to translational building blocks. While the feature and learning pipeline is able to compensate some distortions, strong perspective, scaling, or rotation cannot be handled. We believe that this can be extended quite easily; the main challenge is that we need to replace the global alignment of co-occurrence maps with local alignments that take the coordinate frame of pairwise matches into account. Second, as shown in the pressure test, statistics favors the grid-based detection when the demand of shape accuracy is high. This implies higher level regularity, when observed, should be incorporated to improve our detection. This is important because only high quality detection is able to help later applications. For example, inaccurate detection can only downgrades image synthesis by imposing useless constraints. Further, although we have tried to build tiling grammar using building blocks as discrete elements, the inference of such grammar is combinatorially hard. In many cases our found little or no improvement from the grammar. Here, we could think of including more comprehensive, global relations such as symmetry or hierarchy~\cite{Hu2013PPI}, or we could use the inverse procedural modeling approach in the spirit of \cite{BOKELOHsig2010,Sylvain2010AT} to obtain stronger guarantees for a subset of structures. In summary, conventional pixel-level synthesis has to rely on MRF-based texture generation; while guidance helps, we still observe artifacts and failure cases. A more fundamental understanding of image structure would be necessary to obtain more powerful models.
